---
title: "Hakimi_400Lab1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Please make sure these libraries are loaded
```{r}
library(tidyverse)
library(data.table)

```

Make sure the web data is in your working directory
```{r Problem 1 Load}

web.dt <- fread("webtraffic.txt")

```


Problem #1
a)
```{r 1a}
traffic <- matrix(colSums(web.dt), nrow = 9, ncol = 9, byrow = TRUE)

colnames(traffic) <- 1:9
rownames(traffic) <- 1:9
traffic[9,1] <- 1000
print(traffic)

```
b)
The Markov Chain is both irreducable and egrodic because all states "communicate"
with each other in some chain and becuase

c)
```{r 1c}
P <- traffic/rowSums(traffic)
print(P)
```
d)
```{r 1d}
init <- c(1, rep(0,8))
prob5 <- P %*% P %*% P %*% P %*% P
print(prob5[1,5])

```
e)
```{r 1e}
library(expm)
ssprobmat <- P %^% 10000000
print(ssprobmat[1,])

```

f)
```{r 1f} 
ident = diag(9)
Q <- ident[c(1:8), c(1:8)]- P[c(1:8), c(1:8)]
Qin <- solve(Q)
mins <- c(0.1,2,3,5,5,3,3,2)
ans <- Qin %*% mins
print('Avg. time spent on site (MINUTES):')
print(ans[1])

```

Problem 2

a)
``` {r 2a}
lam1 = 1
lam2 = 2
lam3 = 4

var_px = 1/lam1^2
tolerance = 1 * 10^-3
delta = 1 - .99
n1 = var_px/(tolerance^2 * delta) 

var_px = 1/lam2^2
tolerance = 1 * 10^-3
delta = 1 - .99
n2 = var_px/(tolerance^2 * delta) 

var_px = 1/lam3^2
tolerance = 1 * 10^-3
delta = 1 - .99
n3 = var_px/(tolerance^2 * delta) 

print('Samples needed for lambda = 1') 
print(n1)
print('Samples needed for lambda = 2') 
print(n2)
print('Samples needed for lambda = 4') 
print(n3)

```

b)
``` {r 2b}
# lambda 1
X = runif(n1, 0,1)
Y = -log(X)/lam1
g = sin(Y)/lam1
print('Estimation with Lambda 1')
print(sum(g)/n1)
print('Actual Value with Lambda 1')
print(1/(1+lam1^2))

# lambda 2
X = runif(n2, 0,1)
Y = -log(X)/lam2
g = sin(Y)/lam2
print('Estimation with Lambda 2')
print(sum(g)/n2)
print('Actual Value with Lambda 2')
print(1/(1+lam2^2))

# lambda 3
X = runif(n3, 0,1)
Y = -log(X)/lam3
g = sin(Y)/lam3
print('Estimation with Lambda 3')
print(sum(g)/n3)
print('Actual Value with Lambda 3')
print(1/(1+lam3^2))

```
All values are within the 1*10^-3 error margin thanks to the number of samples


Problem 3

a)
Metropolis-Hastings is best suited for this problem. 

b)
```{r 3b}
q = function(x){
  return(rexp(1, rate = x))
}

p = function(x){
  return(dgamma(x, shape = 1, rate = 1))
}


x = rep(0,15000)
x[1] = 3     #initialize; I've set arbitrarily set this to 3
for(i in 2:15000){
  current_x = seq(i-1)
  proposed_x = q(current_x)
  A = min(1,(p(proposed_x) * q(current_x)) / (p(current_x) * q(proposed_x)))
  if(runif(1)<=A){
    x[i] = proposed_x       # accept move with probabily min(1,A)
  } else {
    x[i] = current_x        # otherwise "reject" move, and stay where we are
  }
}


hist(x, breaks = 10, prob = TRUE)
curve(dgamma(x, shape = 2, scale = 2), add=TRUE,col="red")


```

c)
```{r 3c}

plot(x)

```
Yes this graph shows that the points generated in part 3b
are sufficuently random.

